Comparison of particle trajectories and collision operators for collisional transport in nonaxisymmetric plasmas

### 2014-02-04 doc/sfincsPaper/sfincsPaper.tex
Added appendix on Simakov benchmark

A similar difference between the momentum-conserving and
pitch-angle-scattering operators is evident in the parallel
conductivity coefficient $L_{33}$. The flow that arises in response to
a parallel electric field is determined by the parallel momentum equation
\begin{equation} 
n_a Z_a e E_{||} = \vect{b}\cdot \nabla p_a + \vect{b}\cdot(\nabla \cdot \vect{\pi}_a )+ F_{a||}, 
\end{equation} 
where the parallel friction force $F_{\|a}$ again vanishes when only a
single ion species is considered. Hence 
\begin{equation} 
n_a Z_a e \left< B  E_{||} \right> = \left< \vect{B}\cdot(\nabla \cdot \vect{\pi}_a) \right> 
= \left< (p_\perp - p_\| ) \nabla_\| B \right>.
\end{equation}
In
the absence of radial gradients, the pressure anisotropy in the \PS
regime is proportional to the parallel flow velocity and inversely
proportional to the collision frequency \cite{AndreiPer2009}. The
flow $\left< V_\| B \right>$ is therefore proportional to $\nu$ in
the \PS regime unless momentum conservation is violated. In the
latter case, the spurious friction force causes $\left< V_\| B
\right>$ to be inversely proportional to $\nu$, as can be seen in
Figs.~(\ref{fig:collisionComparison_LHD})-(\ref{fig:collisionComparison_W7X}).

@@

>>>>

A similar difference between the momentum-conserving and
pitch-angle-scattering operators is evident in the parallel
conductivity coefficient $L_{33}$. The flow that arises in response to
a parallel electric field is determined by the parallel momentum equation
\begin{equation} 
n_a Z_a e E_{||} = \vect{b}\cdot \nabla p_a + \vect{b}\cdot(\nabla \cdot \vect{\pi}_a )+ F_{a||}, 
\end{equation} 
where the parallel friction force $F_{\|a}$ again vanishes when only a
single ion species is considered. Hence 
\begin{equation} 
n_a Z_a e \left< B  E_{||} \right> = \left< \vect{B}\cdot(\nabla \cdot \vect{\pi}_a) \right> 
= \left< (p_\perp - p_\| ) \nabla_\| B \right>.
\label{ParallelMomentum}
\end{equation}
In
the absence of radial gradients, the pressure anisotropy in the \PS
regime is proportional to the parallel flow velocity and inversely
proportional to the collision frequency \cite{AndreiPer2009}. The
flow $\left< V_\| B \right>$ is therefore proportional to $\nu$ in
the \PS regime unless momentum conservation is violated. In the
latter case, the spurious friction force causes $\left< V_\| B
\right>$ to be inversely proportional to $\nu$, as can be seen in
Figs.~(\ref{fig:collisionComparison_LHD})-(\ref{fig:collisionComparison_W7X}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%Appendix on Simakov benchmark

\section{Ion transport matrix at high collisionality}
From the model presented in Ref.~\cite{AndreiPer2009} we can derive expressions for $L_{jk}$ of Eq.~(\ref{eq:matrix}) in the Pfirsch-Schl{\"u}ter regime. 
The coefficients then depend on the function $u$ given by the solution to Eq.~(8) in Ref.~\cite{AndreiPer2009}, which is proportional to the parallel current divided by $B$. 
All coefficients but $L_{33}$ are straightforwardly obtained from Eqs.~(14), (18) and (26) in Ref.~\cite{AndreiPer2009} for the radial ion heat flux, the parallel ion flow and the radial current respectively, by %properly choosing 
suitable choices of 
the thermodynamic forces in the right-hand-side of Eq.~(\ref{eq:matrix}) for different solves and using the symmetry of the transport matrix.
To find the parallel conductivity coefficient $L_{33}$ we substitute the pressure anisotropy, given by Eq.~(20) in Ref.~\cite{AndreiPer2009}, into Eq.~(\ref{ParallelMomentum}) (of this paper) 
in the absence of radial gradients (i.e. when $E_{||}$ is the only thermodynamic force present). $L_{33}$ is then found from the flow $\left< V_\| B \right>$.

Expressions for the matrix coefficients in the Pfirsch-Schl{\"u}ter regime are summarized in Eqs.~(\ref{eq:LPSregime})-(\ref{eq:PSregimeFunctions}). 
Note that the constants in the function ${K}_2 \left(\psi\right)$ in $L_{23}$ have been modified from their values in ${K}_{2}^{\mathrm{Simakov}} \left(\psi\right)$ obtained from Ref.~\cite{AndreiPer2009}, by a more precise solution to the Spitzer problem.

%\begin{equation}
%\begin{array}{c}
\begin{align}
\begin{split}
L_{11} & = 
 0.96 \cdot 2^{1/2} \cdot  \frac{3}{4}  \frac{\left(G + \iota I\right)^2}{ \iota^2 G^2} 
 {G}_1 \left(\psi\right) \frac{1}{\nu'},\\
 %%
L_{12} & = L_{21} =  0.96 \cdot 2^{1/2} \frac{\left({G} + \iota {I}\right)^2}{ \iota^2 {G}^2} \left[ 3.245 \, {G}_1 \left(\psi\right) + 0.085 \, {G}_2 \left(\psi\right)  \right] \frac{1}{\nu'}, \\
%% 
L_{13} & = L_{31}  = \frac{\left\langle {u} {B}^2  \right\rangle}{{G} \iota} - \frac{\left\langle {B}^2\right\rangle}{{G} \iota} {K}_1\left(\psi\right), \\
%%
L_{22} & = 2^{1/2} \cdot  \frac{8}{5} \frac{1}{\iota^2 {G}^2} B_0^2 {H} \left(\psi\right) \, \nu', \\
	%\textcolor{red}{+ \left( \,0.96 \cdot 2^{1/2} \cdot \frac{5}{2} \frac{\left({G} + \iota {I}\right)^2}{ \iota^2 {G}^2} \left[ 3.245 {G}_1 \left(\psi\right) + 0.085 {G}_2 \left(\psi\right)  \right] \frac{1}{\nu'}\right)}, \\
	%%
L_{23} & = L_{32} = \frac{1}{\iota {G}} \left[\frac{5}{2} \left\langle {u} {B}^2  \right\rangle - \frac{5}{2} {K}_1\left(\psi\right) \left\langle {B}^2\right\rangle + {K}_2\left(\psi\right) \left\langle {B}^2\right\rangle\right], \\
%%
L_{33} & = \frac{1}{3 \cdot 0.96 \cdot 2^{1/2}} \frac{1}{\left({G} + \iota {I}\right)^2} \frac{\left\langle {B}^2\right\rangle^2}{\left\langle \left( {\nabla}_\| {B}\right)^2\right\rangle} \, \nu',	
%\end{array}
\label{eq:LPSregime}
\end{split}
\end{align}
%\end{equation}
%where
\begin{align}
\begin{split}
& {G}_1 \left(\psi\right) %& 
= \frac{\left\langle \left( {\nabla}_\| \ln {B}\right)  {\nabla}_\| \left({u} {B}^2\right)\right\rangle^2}{\left\langle \left( {\nabla}_\| {B}\right)^2\right\rangle} -
 \left\langle \left[\frac{ {\nabla}_\| \left({u} {B}^2\right)}{{B}}\right]^2\right\rangle,  \\
 %%
& {G}_2 \left(\psi\right) %& 
 = \left\langle {u} \left({\nabla}_\| \ln {B}\right) {\nabla}_\| \left({u} {B}^2\right) \right\rangle
 - \frac{\left\langle \left( {\nabla}_\| \ln {B}\right)  {\nabla}_\| \left({u} {B}^2\right)\right\rangle
 \left\langle {u} \left( {\nabla}_\| {B}\right)^2\right\rangle}{\left\langle \left( {\nabla}_\| {B}\right)^2\right\rangle},  \\
 %%
& {K}_1 \left(\psi\right) %& 
 = \frac{\left\langle \left( {\nabla}_\| \ln {B}\right)  {\nabla}_\| \left({u} {B}^2\right)\right\rangle}{2 \left\langle \left( {\nabla}_\| {B}\right)^2\right\rangle}, \\
 %%
& {K}_2 \left(\psi\right) %& 
 = 1.97213 \frac{\left\langle {u} {B}^2  \right\rangle}{\left\langle {B}^2  \right\rangle} - 1.03287 \cdot 2 {K}_1 \left(\psi\right)
 + 0.09361 \frac{
 \left\langle {u} \left( {\nabla}_\| {B}\right)^2\right\rangle}{\left\langle \left( {\nabla}_\| {B}\right)^2\right\rangle}, \\
 %%
&  {K}_{2}^{\mathrm{Simakov}} \left(\psi\right) %& 
  = 1.77 \frac{\left\langle {u} {B}^2  \right\rangle}{\left\langle {B}^2  \right\rangle} - 0.91 \cdot 2 {K}_1 \left(\psi\right)
 + 0.05 \frac{
 \left\langle {u} \left( {\nabla}_\| {B}\right)^2\right\rangle}{\left\langle \left( {\nabla}_\| {B}\right)^2\right\rangle}, \\
 %%
& {H} \left(\psi\right) %& 
 = \frac{\left\langle {u} {B}^2  \right\rangle^2}{\left\langle {B}^2  \right\rangle} - \left\langle {u}^2 {B}^2  \right\rangle.
\label{eq:PSregimeFunctions}
\end{split}
\end{align}
